---
title: "open-data-movilidad"
output: 
  github_document:
    html_preview: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE)
library(tidyverse)
library(sf)
library(tmap)
library(kableExtra)
tmap_mode("view")
sf::sf_use_s2(FALSE)
```


This repository contains code to access OD data (based on mobile phones locations) from the Ministry of Transport, Mobility and Urban Agenda (Spain). 

The raw data and some information on the method used to collect the data can be found in the link below. (Do not use Chromium to open the link or you won't be able to see the data!)

If this data is periodically updated, it may be worth it to create a data r package.

https://www.mitma.gob.es/ministerio/covid-19/evolucion-movilidad-big-data/opendata-movilidad 


```{r}
# Read the data
dist_202102 = read.table(gzfile("datos/202102_maestra1_mitma_distrito/20210201_maestra_1_mitma_distrito.txt.gz"), header = TRUE, sep = "|")
head(dist_202102)
# dist_202102_2 = read.table(gzfile("datos/202102_maestra2_mitma_distrito/20210201_maestra_2_mitma_distrito.txt.gz"), header = TRUE, sep = "|")
# names(dist_202102_2)
# muni_202102 = read.table(gzfile("datos/202102_maestra1_mitma_municipio/20210201_maestra_1_mitma_municipio.txt.gz"), header = TRUE, sep = "|")
# names(muni_202102)
# muni_202102_2 = read.table(gzfile("datos/202102_maestra2_mitma_municipio/20210201_maestra_2_mitma_municipio.txt.gz"), header = TRUE, sep = "|")
# names(muni_202102_2)

dist = read.table("relaciones_distrito_mitma.csv", header = TRUE, sep = "|")
# muni = read.table("relaciones_municipio_mitma.csv", header = TRUE, sep = "|")

zonificacion_dist <- sf::st_read(
  "zonificacion_distritos/zonificacion-distritos/distritos_mitma.shp")
# zonificacion_muni <- sf::st_read(
#   "zonificacion_municipios/zonificacion-municipios/municipios_mitma.shp")

cod_pro = read.table("cod_ccaa_pro.csv", header = TRUE, sep = ",")
```

```{r}
# create province and district origin
dist_202102 = dist_202102 %>% 
  # slice_head(n = 10) %>%
  separate(origen, into = c("prov_O", "dist_O"), sep = 2, remove = FALSE)  

# filter potetial cyclable trips (< 10km) and group by district
dist_202102 = dist_202102 %>% 
  filter(distancia == "002-005" | distancia == "005-010") %>% 
  group_by(origen, prov_O, dist_O) %>% 
  summarise("num. trips <10 km"= sum(viajes))

# add geometry and province and region tags
dist_202102 = dist_202102 %>% 
  left_join(zonificacion_dist, by = c("origen" = "ID"))
dist_202102$prov_O = as.numeric(dist_202102$prov_O)
dist_202102 = dist_202102 %>% 
  left_join(cod_pro, by = c("prov_O" = "Cod_Provincia"))

table(dist_202102$Nom_CCAA)

# make map plottable
dist_202102 = st_sf(dist_202102)
```

```{r}
# filter Valencian Country
dist_202102_CV = dist_202102 %>% 
   filter(Nom_CCAA == "COMUNIDAD VALENCIANA")

# plot map
qtm(dist_202102_CV, "num. trips <10 km")
```

*Fig 1: Number of trips under 10 km by district (during February 2021) Valencian Country*

